---
title: "Week1"
author: "Jinyoung Seo"
date: "3/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Week 1

```{r}
getwd()
dir()
#read.csv("mydata.csv")

```
- Get working directory and files in the working directory
```{r}
#myfunction()

source("mycode.R")
ls()
objects()
```

- ls and objects gives the list of objects in the environment


```{r}
second(4)


remove(a)


``` 

### Entering Input
```{r}

x <-1
print(x)
x
msg <-"hello"
```


### Printing 
```{r}
x<- 1:5
x
```
The : operator is used to create integer sequences


### Objects
R has five basic or 'atomic' objects
- character
- numeric
- integer
- complex
- logical (True/False)

The most basic object is a vector
- A vector can only contain objects of the same class
- But: the one exception is a list, which is represented as a vector but can contain objects of different classes 

Empty vector can be created with the `vector()` function


### Numbers
- If you explicitly want an integer, you need to specify the L suffix
- EX: 1L gives you an integer


### Attributes
R objects can have attributes
- names, dimnames
- dimensions (e.g. matrices, arrays)
- class
- length
- otehr user-defined attributes/metadata

Attributes of an objects can be accessed using the `attributes` function


### Creating vectors
`c()` function can be used to create vectors of objects

```{r}
x <- c(1,2)         ## numeric
x <- c(TRUE,FALSE)  ## logical
x <- c(T,F)         ## logical
x <- c("a","b","C") ## character
x <- 9:29           ## Integer
x <- c(1+0i,2+4i)   ## complex
```

using `vector()` function
```{r}
x<- vector("numeric",length=10)
```


### Explicit coercion
Objects can be explicitly coerced from one to another using `as.*` functions

```{r}
x<-0:6
class(x)
as.numeric(x)
as.logical(x)
as.character(x)
```


### List
```{r}
x<-list(1,"a",TRUE,1+4i)
x
```



### Matrices
Matrices are constructed column-wise
```{r}
m <- matrix(nrow=2,ncol=3)
m
dim(m)
attributes(m)
```

Matrices can also be created directly from vectors by adding a dimension attribute
```{r}
m <-1:10
m
dim(m) <- c(2,5)
m
```


### cbind-ing and rbind-ing
```{r}
x <- 1:3
y <- 10:12
cbind(x,y)
rbind(x,y)
```


### Factors
Factors are used to represent categorical data. Factors can be unordered or ordered. One can think of a factor as an integer vector where each integer has a *label*

- Factors are treated specially by modelling functions like `lm()` and `glm()`
- Using factors with labels is better than using integers because factors are self-describing; having a variable that has values "Male" or "Female" is better than a variable that has values 1 and 2
```{r}
x <-factor(c("yes","yes","no","yes","no"))
x
table(x)
unclass(x)
```

It is also possible to specify which is the base level. Unless specified explicitly, levels follow alphabetical order
```{r}
x <-factor(c("yes","yes","no","yes","no"))
x
x <-factor(c("yes","yes","no","yes","no"),levels=c("yes","no"))
x
```

### Missing values
- NAN is used for undefined mathematical operations
- `is.nan()` tests objects if they are NAN
- NA is for everything else
- `is.na()` tests objects if they are NA
- NAN is also considered as NA



### Data frames
- Data frames are used to stroe tabular data in R
- Data frames are represented as a special type of list where every element of the list has to have the same length
- Each element of the list can be thought of as a column and the length of each element of the list is the number of rows
- Unlike matrices, data frames can store different classes of objects in each column
- Data frames also have a special attribute called `row.names`
- Data frames are usually created by calling `read.table()` or `read.csv()`
- Can be converted to a matrix by calling `data.matrix()`

```{r}
x<- data.frame(foo=1:4,bar=c(T,F,T,T))
x
nrow(x)
ncol(x)
```

### Names
R objects can also have names
```{r}
x<- 1:3
names(x)
names(x)<-c("foo","bar","norf")
names(x)
x<-list(a=1,b=2,c=3)
x
x<-matrix(1:4,nrow=2,ncol=2) ## First element of the list is row names
dimnames(x)<-list(c("a","b"),c("c","d"))
x
```


### Reading data
- `read.table, read.csv` for reading tabular data -> data frame
- `readLines` for reading lines of a text file
- `source` for reading in R code files (inverse to `dump`)
- `dget` for reading in R code files (inverse to `dput`)
- `load` for reading in saved workspaces
- `unserialize` for reading single R objects in binary form


### Write data
- `write.table` for writing tabular data to text files (csv) or connections
- `writeLines` for writing character data line-by-line to a file or connection
- `dump` for dumping a textual representation of multiple R objects
- `dput` for outputting a textual representation of an R objects
- `save` ofr saving an arbitrary number of R objects in binary format to a file
- `serialize` for converting an R object into a binary format for outputting to a connection


### Reading in larger datasets with read.table
- Set `comment.char=""` if there are no commented lines in your file
- Use the `colClasses` argument. Specifying this option instead of using the default can make `read.table` run **much** faster
- In order to use this option, you have to know the class of each column in your data frame
  - If all of the columns are "numeric" then you can just set `colClasses="numeric"`
  - A quick and dirty way to figure out the classes of each column is the following
  ```{r, eval=FALSE, tidy=TRUE}
  initial <- read.table("datatable.txt",nrows=100)
  classes <- sapply(initial,class)
  tabAll  <- read.table("datatable.txt",colClasses=classes)
  ```


```{r}

y <- data.frame(a=1,b="a")
y
dput(y)
dput(y,file="y.R")
new.y <- dget("y.R")
```

### Subsetting R objects
- [ operator always return an object of the same class as the original. It can be used to select multiple elements of an object
- [[ operator is used to extract elements of a list or a data frame. It can only be used to extract a single element and the class of the returned object will not necessarily be a list or data frame
- $ operator is used to extract elements of a list or data frame by literal name. Its semantics are similar to that of [[

```{r}
x <- c("a","b","c","c","d","a")
x[1]
x[2]
x[1:4]
x[c(1,3,4)]
u <- x>"a"
u
x[u]
x[x>"a"]
```

### Subsetting lists
```{r}
x <- list(foo=1:4, bar="0.6")
x
x[[1]]
x[[2]]
x[["bar"]]
x$bar
```
### Subsetting nested elements of a list
```{r}
x<- list(a=list(10,12,14),b=c(3.14,2.81))
# Get the 3rd element of the 1st element
x[[c(1,3)]]
x[[1]][[3]]
x[[c(2,1)]]
# [ operator can be used to extract multiple elements from a list.
x[c(1,2)]
```

### Subsetting a matrix
```{r}
x <- matrix(1:6, 2, 3)
x[1,2]
x[1,]
# Drop matrix dimension
x[1,2,drop=TRUE] # Hence, it's a vector of length 1
# Preserve dimension
x[1,2,drop=FALSE] # Hence, it's a 1x1 matrix
x[1, ]
x[1, ,drop=FALSE]
read.csv("hw1_data.csv")
data[data[,1]>31,][data[,4]>90,]
mean(data[data[,"Month"]==6,][,"Temp"],na.rm = TRUE)
```

### Partial matching
```{r}
x <- list(aardvark = 1:5,abb=6:7)
x$aa
x[["aa"]]
x[["aa",exact=FALSE]]
```


## Removing NA values
```{r}
x <- c(1, 2, NA, 4, NA, 5)
bad <- is.na(x)
print(bad)
x[!bad]
```
What if there are multiple R objects and you want to take the subset with no missing values in any of those objects?
```{r}
x <- c(1, 2, NA, 4, NA, 5)
y <- c("a", NA, NA, "d", NA, "f")
good <- complete.cases(x,y)
good
x[good]
y[good]
```
You can use `complete.cases` on data frames too.
```{r}
head(airquality,n=8)
good <- complete.cases(airquality)
head(airquality[good,])
```